Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-05 troll/supp_body.tex
first pass on troll appendix

\section{Proofs from Section~\ref{s:algbatch}}
\subsection{Proof of Theorem~\ref{t:active}}
The following ancillary results will be useful.
%
\begin{lemma}[Hoeffding's inequality for sampling without replacement]
\label{l:hoeff-wr}
Let $\scX = \{x_1,\dots,x_N\}$ be a finite subset of $[0,1]$ and let
\[
	\mu = \frac{1}{N}\sum_{i=1}^N x_i~.
\]
If $X_1,\dots,X_n$ is a random sample drawn at random from $\scX$ without replacement, then, for every $\ve > 0$,
\[
	\Pr\left( \left|\frac{1}{n}\sum_{t=1}^n X_t - \mu \right| \ge \ve \right) \le 2e^{-2n\ve^2}~.
\]
\end{lemma}
%
\begin{lemma}\label{l:bins}
% Given a function $f\,:\, \Nat \rightarrow \R$ such that $f(n) \rightarrow \infty$ as $n \rightarrow \infty$, 
Let $\NN_1, \ldots, \NN_n$ be subsets of a finite set $E$.
% with $|E| \geq n\,f(n)$
% Let $\theta = \frac{|E|}{n\,f(n)}$, 
% and $k = k(\theta)$ be the number of subsets $\NN_i$ such that $|\NN_i| \geq \theta$. Let 
Let $E_0 \subseteq E$ be sampled uniformly at random {\em without} replacement from $E$, with $|E_0| = m$. Then, for $\delta \in (0,1)$, $Q > 0$, and $\theta \geq 2\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$, we have
\[
\Pr\Bigl( \exists i \,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\Bigl) \leq \delta
\]
provided $|E| \geq m \geq \frac{2|E|}{\theta}\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$.
\end{lemma}
%
\begin{proof}[Proof of Lemma \ref{l:bins}]
Set for brevity $p_i = |\NN_i|/|E|$. Then, due to the sampling without replacement, each random variable $|\NN_i \cap E_0|$ is the sum of $m$ dependent Bernoulli random variables $X_{i,1}, \ldots, X_{i,m}$ such that $\Pr(X_{i,t} = 1) = p_i$, for $t = 1, \ldots, m$. Let $i$ be such that $|\NN_i| \geq \theta$. Then the condition $m \geq \frac{2|E|Q}{\theta}$ implies
\[
Q \leq \frac{m\theta}{2|E|} \leq \frac{m\,p_i}{2} = \frac{\E\big[|\NN_i \cap E_0|\big]}{2}~.
\]
Since the variables $X_{i,j}$ are negatively associated, we may apply a (multiplicative) Chernoff bound~\cite[Section~3.1]{dpbook}.
This gives
\[
\Pr\big(|\NN_i \cap E_0| < Q\big) \leq e^{-\frac{m\,p_i}{8}} \leq e^{-\frac{m\,\theta}{8|E|}}
\]
so that
$
\Pr\bigl( \exists i\,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\bigr) \leq n\,e^{-\frac{m\,\theta}{8|E|}}
$,
which is in turn upper bounded by $\delta$ whenever 
$m \geq \frac{8|E|}{\theta}\ln \frac{n}{\delta}$. % = 8\,n\,f(n)\,\ln \frac{n}{\delta}$.
\end{proof}

Let now $E_{\theta} = \theset{(i,j) \in E}{\din(j) \ge \theta,\, \dout(i) \ge \theta} \setminus E_0$, where $E_0 \ss E$ is the set of edges sampled by the learning algorithm of Section~\ref{s:algbatch}. Then Theorem~\ref{t:active} in the main paper is an immediate consequence of the following lemma. 
%
\begin{lemma}\label{l:active}
Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model in Section~\ref{s:gen}.
For all $0 < \alpha,\delta < 1$ and $0 < \ve < \tfrac{1}{16}$, if the learning algorithm of Section~\ref{s:algbatch} is run with parameters $Q = \tfrac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ and $\alpha$, then with probability at least $1-11\delta$ the predictions $\yhat(i,j)$ satisfy $\yhat(i,j) = y^*(i,j)$ for all $(i,j) \in E_{\theta}$ such that $\big|\eta(i,j) - \tfrac{1}{2}\big| > 8\ve$.
\end{lemma}
%
\begin{proof}[Proof of Lemma \ref{l:active}]
We apply Lemma~\ref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\ln \frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and $\Nin(i),\Nout(i)$, for $i \in V$. We have that, with probability at least $1-\delta$, at least $Q$ edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(i)$ are sampled for each $i$ such that $\NNin(i) \ge \theta$, and at least $Q$ edges of $\Nout(j)$ are sampled for each $j$ such that $\NNout(j) \ge \theta$.
For all $(i,j) \in E_{\theta}$ let
	\pbar_j = \frac{1}{\din(j)}\sum_{i\in\NNin(j)} p_i \quad\text{and}\quad \qbar_i = \frac{1}{\dout(i)}\sum_{j\in\NNout(i)} q_j
and set for brevity $\hdeltain(j) = 1-\hun(j)$ and $\hdeltaout(i) = 1-\htr(i)$.
We now prove that $\hdeltain(j)$ and $\hdeltaout(i)$ are concentrated around their expectations for all $(i,j) \in E_{\theta}$. Consider $\hdeltaout(i)$ (the same argument works for $\hdeltain(j)$). Let $J_1,\dots,J_Q$ be the first $Q$ draws in $E_0 \cap \NNout(i)$ and define
	\muhat_p(i) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_i+q_{J_t}}{2}~.
Applying Lemma~\ref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(i)}$, and using our choice of $Q$, we get that $\big|\muhat_p(i) - \mu_p(i)\big| \le \ve$ holds with probability at least $1-\delta/(2|V|)$, where
	\mu_p(i) = \frac{1}{\dout(i)}\sum_{j \in \NNout(i)} \frac{p_i + q_j}{2} = \frac{p_i + \qbar_i}{2}~.
Now consider the random variables $Z_t = \Ind{y_{i,J_t}=1}$, for $t=1,\dots,Q$. Conditioned on $J_1,\dots,J_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid J_t] = \tfrac{p_i+q_{J_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and using our choice of $Q$, we get that
	\left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(i) \right| \le \ve 
with probability at least $1-\delta/(2|V|)$ for every realization of $J_1,\dots,J_Q$. 
%
Since $\hdeltaout(i) = (Z_1+\cdots+Z_Q)/Q$, we get that $\big|\hdeltaout(i) - \mu_p(i)\big| \le 2\ve$ with probability at least $1-2\delta/(2|V|)$. Applying the same argument to $\hdeltain(j)$, and the union bound\footnote
{
The sample spaces for the ingoing and outgoing edges of the vertices occurring as endpoints in $E_{\theta}$ overlap. Hence, in order to prove a uniform concentration result, we need to apply the union bound over the random variables defined over these sample spaces, which motivates the presence of the factor $\ln(2|V|)$ in the definition of $Q$. 
}
on the set $\theset{\hdeltain(j),\hdeltaout(i)}{(i,j) \in E_{\theta}}$, we get that

@@

	\left| \hdeltaout(i) + \hdeltain(j) - \frac{p_i + q_j}{2} - \frac{\pbar_j + \qbar_i}{2} \right| \le 4\ve

@@

simultaneously holds for all $(i,j) \in E_{\theta}$ with probability at least $1 - 4\delta$. Now notice that $\pbar_j$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$. Similarly, $\qbar_i$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1 \mu\big(p,\cdot\big) dp$ with expectation $\mu_q$. By applying Hoeffding bound for independent variables, together with the union bound to the set of pairs of random variables whose sample means are $\pbar_j$ and $\qbar_i$ for each $(i,j) \in E_{\theta}$ (there are at most $2|V|$ of them) we obtain that

@@

hold simultaneously for all $(i,j) \in E_{\theta}$ with probability at least $1-2\delta$. Combining with~(\ref{eq:first}) we obtain that

@@

	\left| \hdeltaout(i) + \hdeltain(j) - \frac{p_i + q_j}{2} - \frac{\mu_p + \mu_q}{2} \right| \le 5\ve

@@

simultaneously holds for each $(i,j) \in E_{\theta}$ with probability at least $1 - 6\delta$. Next, let $E_L'$ be the set of the first $Q$ edges drawn in $E_L \cap E_0$. Then

@@

	\E\big[\tauhat\big] = \frac{1}{Q}\sum_{(i,j) \in E_L'} \Pr\big(y_{i,j} = 1\big) = \frac{1}{Q}\sum_{(i,j) \in E_L'} \frac{p_i+q_j}{2}~,

@@

where the expectation is w.r.t.\ the independent draws of the labels $y_{i,j}$ for $(i,j) \in E_L'$. Hence, by applying again Hoeffding bound (this time without the union bound) to the $Q = \tfrac{1}{2\ve^2}\ln\tfrac{2}{\delta}$ independent Bernoulli random variables $\Ind{y_{i,j} = 1}$, $(i,j) \in E_L'$, the event

@@

	F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(i,j) \in E_L'} \frac{p_i+q_j}{2}~.

@@

For any realization $\bq_0$ of $\bq$, the function $F_1(\bp) = F(\bp,\bq_0)$ is a sample mean of $Q = \tfrac{1}{2\ve^2}\ln\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables $\theset{p_i}{(i,j) \in E_L'}$ (recall that if $i \in V$ is the origin of an edge $(i,j) \in E_L'$, then it is not the origin of any other edge $(i,j') \in E_L'$). Using again the standard Hoeffding bound, we obtain that

@@

holds with probability at least $1 -\delta$ for each $\bq\in [0,1]^{|V|}$. With a similar argument, we obtain that

@@

	\left| \hdeltaout(i) + \hdeltain(j) - \tauhat - \frac{p(i) + q(j)}{2} \right| \le 8\ve

@@

simultaneously holds for each $(i,j) \in E_{\theta}$ with probability at least $1 - 10\delta$. Putting together concludes the proof.

@@

%

\iffalse
****************************************************
\begin{lemma}[Hoeffding's inequality]
\label{l:hoeff}
Let $X_1,\dots,X_n$ be independent random variables such that $X_i \in [0,1]$ and let
\[
	s_n = \frac{1}{n}\sum_{t=1}^n \big(X_t - \E[X_t]\big)~.
\]
Then, for every $\ve > 0$,
$
	\Pr\big(|s_n| \ge \ve\big) \le 2e^{-2n\ve^2}~.
$
\end{lemma}
%
****************************************************
\fi

\subsection{Derivation of the maximum likelihood equations}
%
Recall that the training set
$
	E_0 = \theset{\big(i_t,j_t),y_{i_t,j_t}\big)}{t=1,\dots,m}
$
is drawn uniformly at random from $E$ without replacement. 
%Set for brevity $m = |E_0|$. 
We can write
%
\begin{align*}
	\Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)\\ 
&= 
	\frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{i_k}+q_{j_k}}{2} \right)^{\Ind{y_{i_k,j_k}=+1}}\,\\
&\quad\times\,\prod_{k=1}^m\left(1-\frac{p_{i_k}+q_{j_k}}{2} \right)^{\Ind{y_{i_k,j_k}=-1}}\\
&=
\frac{1}{\binom{|E|}{m}\,m!}
\prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
\left(\frac{p_{\ell}+q_{j_k}}{2} \right)^{\Ind{i_k = \ell,\,y_{\ell,j_k}=+1}}\,\\
&\quad\times\,
\prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{j_k}}{2} \right)^{\Ind{i_k = \ell,\,y_{\ell,j_k}=-1}}\Biggl)
\end{align*}
%
so that $\log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)$ is proportional to
%
\begin{align*}
&\sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=+1}\,
\log\left(\frac{p_{\ell}+q_{j_k}}{2} \right)\,\\
&+\,
\sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=+1}\, 
\log\left(1-\frac{p_{\ell}+q_{j_k}}{2} \right)
\end{align*}
%
and
%
\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
&= 
\sum_{k=1}^m 
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\,\\
& - \,
\sum_{k=1}^m \frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}}\,.
\end{align*}
%
By a similar argument,
%
\begin{align*}
\Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
\\ &=
\frac{1}{\binom{|E|}{m}\,m!}
\prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
\left(\frac{p_{i_k}+q_{\ell}}{2} \right)^{\Ind{j_k = \ell,\,y_{i_k,\ell}=+1}}\,\\
&\quad\times\,
\prod_{k=1}^m\left(1-\frac{p_{i_k}+q_{\ell}}{2} \right)^{\Ind{j_k = \ell,\,y_{i_k,\ell}=-1}}\Biggl)
\end{align*}
%
so that
%
\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}
&= 
\sum_{k=1}^m 
\frac{\Ind{j_k = \ell,y_{i_k,\ell}=+1}}{p_{i_k}+q_{\ell}}\,\\ 
&- \,
\sum_{k=1}^m 
\frac{\Ind{j_k = \ell,y_{i_k,\ell}=-1}}{2-p_{i_k}-q_{\ell}}\,.
\end{align*}
%
We then derive the approximation presented in the main paper. Namely, equating to zero the
gradient of the log likelihood w.r.t $p_\ell$ gives
\begin{align*}
\sum_{k=1}^m 
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\,
 - \,
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}} &= 0
\end{align*}
To simplify the notation, let $a_k = \Ind{i_k = \ell,y_{\ell,j_k}=+1}$, $b_k=\Ind{i_k =
\ell,y_{\ell,j_k}=+1}$ and $c_k = p_{\ell} + q_{j_k}$, we can rewrite the previous equation as 
\begin{align*}
\sum_{k=1}^m \frac{a_k}{c_k} - \frac{b_k}{2-c_k} &= 0 \\
\sum_{k=1}^m  \frac{a_k(2-c_k) - b_k c_k }{ c_k(2-c_k) } &=0
% \prod_{k=1}^m c_k\prod_{k=1}^m b_k\sum_{k=1}^m \frac{a_k}{b_k} &=\prod_{k=1}^m c_k\prod_{k=1}^m b_k \sum_{k=1}^m\frac{d_k}{c_k}
\end{align*}

The approximation consists in assuming that the denominator $c_k(2-c_k)$ is a
constant for all $k$, and can therefore be disregarded.
Moving $b_kc_k$ to the right hand side and returning to the original variables,
it yields the approximate equation presented in the main paper, namely
\begin{align*}
\sum_{k=1}^m &\Ind{i_k = \ell,y_{\ell,j_k}=+1} \left(2-p_{\ell}-q_{j_k}\right)\\
=
&\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=-1}
(p_{\ell}+q_{j_k})%, \quad \ell = 1, \ldots, |V|
\end{align*}

>>>>

\section{Proofs from \autoref{s:algbatch}}

\subsection{Proof of \autoref{t:active}}

The following ancillary results will be useful.

\begin{lemma}[Hoeffding's inequality for sampling without replacement]
\label{l:hoeff-wr}
Let $\scX = \{x_1,\dots,x_N\}$ be a finite subset of $[0,1]$ and let
\[ \mu = \frac{1}{N}\sum_{i=1}^N x_i~.  \]
If $X_1,\dots,X_n$ is a random sample drawn at random from $\scX$ without replacement, then, for
every $\ve > 0$,
\[
  \Pr\left( \left|\frac{1}{n}\sum_{t=1}^n X_t - \mu \right| \ge \ve \right) \le 2e^{-2n\ve^2}~.
\]
\end{lemma}

\begin{lemma}\label{l:bins}
% Given a function $f\,:\, \Nat \rightarrow \R$ such that $f(n) \rightarrow \infty$ as $n \rightarrow \infty$, 
Let $\NN_1, \ldots, \NN_n$ be subsets of a finite set $E$.
% with $|E| \geq n\,f(n)$
% Let $\theta = \frac{|E|}{n\,f(n)}$, 
% and $k = k(\theta)$ be the number of subsets $\NN_i$ such that $|\NN_i| \geq \theta$. Let 
Let $E_0 \subseteq E$ be sampled uniformly at random {\em without} replacement from $E$, with $|E_0| = m$. Then, for $\delta \in (0,1)$, $Q > 0$, and $\theta \geq 2\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$, we have
\[
\Pr\Bigl( \exists i \,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\Bigl) \leq \delta
\]
provided $|E| \geq m \geq \frac{2|E|}{\theta}\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$.
\end{lemma}

\begin{proof}[Proof of Lemma \ref{l:bins}]
Set for brevity $p_i = |\NN_i|/|E|$. Then, due to the sampling without replacement, each random
variable $|\NN_i \cap E_0|$ is the sum of $m$ dependent Bernoulli random variables $X_{i,1}, \ldots,
X_{i,m}$ such that $\Pr(X_{i,t} = 1) = p_i$, for $t = 1, \ldots, m$. Let $i$ be such that $|\NN_i|
\geq \theta$. Then the condition $m \geq \frac{2|E|Q}{\theta}$ implies
\[
Q \leq \frac{m\theta}{2|E|} \leq \frac{m\,p_i}{2} = \frac{\E\big[|\NN_i \cap E_0|\big]}{2}~.
\]
Since the variables $X_{u,v}$ are negatively associated, we may apply a (multiplicative) Chernoff
bound~\autocite[Section~3.1]{dpbook}.
This gives
\[
\Pr\big(|\NN_i \cap E_0| < Q\big) \leq e^{-\frac{m\,p_i}{8}} \leq e^{-\frac{m\,\theta}{8|E|}}
\]
so that
$
\Pr\bigl( \exists i\,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\bigr) \leq n\,e^{-\frac{m\,\theta}{8|E|}}
$,
which is in turn upper bounded by $\delta$ whenever 
$m \geq \frac{8|E|}{\theta}\ln \frac{n}{\delta}$. % = 8\,n\,f(n)\,\ln \frac{n}{\delta}$.
\end{proof}

Let now $E_{\theta} = \theset{(u,v) \in E}{\din(v) \ge \theta,\, \dout(u) \ge \theta} \setminus
E_0$, where $E_0 \ss E$ is the set of edges sampled by the learning algorithm of
\autoref{s:algbatch}. Then \autoref{t:active} in the main paper is an immediate consequence of the
following lemma.

\begin{lemma}\label{l:active}
Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model
in \autoref{s:gen}.
For all $0 < \alpha,\delta < 1$ and $0 < \ve < \tfrac{1}{16}$, if the learning algorithm of
\autoref{s:algbatch} is run with parameters $Q = \tfrac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ and
$\alpha$, then with probability at least $1-11\delta$ the predictions $\yhat(u,v)$ satisfy
$\yhat(u,v) = y^*(u,v)$ for all $(u,v) \in E_{\theta}$ such that $\big|\eta(u,v) - \tfrac{1}{2}\big|
> 8\ve$.
\end{lemma}

\begin{proof}[Proof of Lemma \ref{l:active}]
We apply \autoref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\ln
\frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and
$\Nin(u),\Nout(u)$, for $i \in V$. We have that, with probability at least $1-\delta$, at least $Q$
edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(u)$ are sampled for each $i$ such that
$\NNin(u) \ge \theta$, and at least $Q$ edges of $\Nout(v)$ are sampled for each $j$ such that
$\NNout(v) \ge \theta$.
For all $(u,v) \in E_{\theta}$ let
\[
  \pbar_j = \frac{1}{\din(v)}\sum_{i\in\NNin(v)} p_i \quad\text{and}\quad \qbar_i = \frac{1}{\dout(u)}\sum_{j\in\NNout(u)} q_j
\]
and set for brevity $\hdeltain(v) = 1-\hun(v)$ and $\hdeltaout(u) = 1-\htr(u)$.
We now prove that $\hdeltain(v)$ and $\hdeltaout(u)$ are concentrated around their expectations for
all $(u,v) \in E_{\theta}$. Consider $\hdeltaout(u)$ (the same argument works for $\hdeltain(v)$).
Let $J_1,\dots,J_Q$ be the first $Q$ draws in $E_0 \cap \NNout(u)$ and define
\[
  \muhat_p(u) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_i+q_{J_t}}{2}~.
\]
Applying \autoref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(u)}$, and using
our choice of $Q$, we get that $\big|\muhat_p(u) - \mu_p(u)\big| \le \ve$ holds with probability at
least $1-\delta/(2|V|)$, where
\[
  \mu_p(u) = \frac{1}{\dout(u)}\sum_{j \in \NNout(u)} \frac{p_i + q_j}{2} = \frac{p_i + \qbar_i}{2}~.
\]
Now consider the random variables $Z_t = \Ind{y_{u,v_t}=1}$, for $t=1,\dots,Q$. Conditioned on
$J_1,\dots,J_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid J_t] =
\tfrac{p_i+q_{J_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and
using our choice of $Q$, we get that
\[
  \left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(u) \right| \le \ve 
\]
with probability at least $1-\delta/(2|V|)$ for every realization of $J_1,\dots,J_Q$.

Since $\hdeltaout(u) = (Z_1+\cdots+Z_Q)/Q$, we get that $\big|\hdeltaout(u) - \mu_p(u)\big| \le
2\ve$ with probability at least $1-2\delta/(2|V|)$. Applying the same argument to $\hdeltain(v)$,
and the union bound\footnote{The sample spaces for the ingoing and outgoing edges of the vertices
occurring as endpoints in $E_{\theta}$ overlap. Hence, in order to prove a uniform concentration
result, we need to apply the union bound over the random variables defined over these sample spaces,
which motivates the presence of the factor $\ln(2|V|)$ in the definition of $Q$.}
on the set $\theset{\hdeltain(v),\hdeltaout(u)}{(u,v) \in E_{\theta}}$, we get that

@@

  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\pbar_j + \qbar_i}{2} \right| \le 4\ve

@@

simultaneously holds for all $(u,v) \in E_{\theta}$ with probability at least $1 - 4\delta$. Now
notice that $\pbar_j$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the
prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$. Similarly, $\qbar_i$ is
a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables independently drawn from the prior
marginal $\int_0^1 \mu\big(p,\cdot\big) dp$ with expectation $\mu_q$. By applying Hoeffding bound
for independent variables, together with the union bound to the set of pairs of random variables
whose sample means are $\pbar_j$ and $\qbar_i$ for each $(u,v) \in E_{\theta}$ (there are at most
$2|V|$ of them) we obtain that

@@

hold simultaneously for all $(u,v) \in E_{\theta}$ with probability at least $1-2\delta$. Combining
with~\eqref{eq:first} we obtain that

@@

  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\mu_p + \mu_q}{2} \right| \le 5\ve

@@

simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 6\delta$. Next,
let $E_L'$ be the set of the first $Q$ edges drawn in $E_L \cap E_0$. Then

@@

  \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \Pr\big(y_{u,v} = 1\big) = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~,

@@

where the expectation is w.r.t.\ the independent draws of the labels $y_{u,v}$ for $(u,v) \in E_L'$.
Hence, by applying again Hoeffding bound (this time without the union bound) to the $Q =
\tfrac{1}{2\ve^2}\ln\tfrac{2}{\delta}$ independent Bernoulli random variables $\Ind{y_{u,v} = 1}$,
$(u,v) \in E_L'$, the event

@@

  F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~.

@@

For any realization $\bq_0$ of $\bq$, the function $F_1(\bp) = F(\bp,\bq_0)$ is a sample mean of $Q
= \tfrac{1}{2\ve^2}\ln\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables
$\theset{p_i}{(u,v) \in E_L'}$ (recall that if $i \in V$ is the origin of an edge $(u,v) \in E_L'$,
then it is not the origin of any other edge $(u,v') \in E_L'$). Using again the standard Hoeffding
bound, we obtain that

@@

holds with probability at least $1 -\delta$ for each $\bq\in [0,1]^{|V|}$. With a similar argument,

@@

  \left| \hdeltaout(u) + \hdeltain(v) - \tauhat - \frac{p(u) + q(v)}{2} \right| \le 8\ve

@@

simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 10\delta$.
Putting together concludes the proof.

@@

\subsection{Derivation of the maximum likelihood equations}

Recall that the training set $\trainset = \theset{\big(u_t,v_t),y_{u_t,v_t}\big)}{t=1,\dots,m}$
is drawn uniformly at random from $E$ without replacement. 
We can write

\begin{align*}
  \Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) = 
  \frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=+1}}\,
  \times\,\prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=-1}}\\
  &=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=+1}}\,\times\,
    \prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=-1}}\Biggl)
\end{align*}

so that $\log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)$ is proportional to

\begin{equation*}
  \sum_{\ell=1}^{|V|}\left(\sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1}\,
  \log\left(\frac{p_{\ell}+q_{v_k}}{2} \right)\, +\,
  \sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1}\, 
\log\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)\right)
\end{equation*}

and

\begin{equation*}
  \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
  = \sum_{k=1}^m 
  \frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
   - \,
  \sum_{k=1}^m \frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}}\,.
\end{equation*}

By a similar argument,

\begin{align*}
  \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=+1}}\,
\times\,
    \prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=-1}}\Biggl)
\end{align*}

so that

\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}
= \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
- \, \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,.
\end{align*}

We then derive the approximation presented in the main paper. Namely, equating to zero the
gradient of the log likelihood w.r.t $p_\ell$ gives
\begin{align*}
\sum_{k=1}^m 
\frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
 - \,
\frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}} &= 0
\end{align*}

To simplify the notation, let $a_k = \Ind{u_k = \ell,y_{\ell,v_k}=+1}$, $b_k=\Ind{u_k =
\ell,y_{\ell,v_k}=+1}$ and $c_k = p_{\ell} + q_{v_k}$, we can rewrite the previous equation as 
\begin{align*}
  \sum_{k=1}^m \frac{a_k}{c_k} - \frac{b_k}{2-c_k} = 0 =
  \sum_{k=1}^m  \frac{a_k(2-c_k) - b_k c_k }{ c_k(2-c_k) }
\end{align*}

The approximation consists in assuming that the denominator $c_k(2-c_k)$ is a
constant for all $k$, and can therefore be disregarded.
Moving $b_kc_k$ to the right hand side and returning to the original variables,
it yields the approximate equation presented in the main paper, namely
\begin{align*}
  \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=+1} \left(2-p_{\ell}-q_{v_k}\right)
  = \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=-1}
  (p_{\ell}+q_{v_k})
\end{align*}

By applying a similar process to the gradient w.r.t. $q_\ell$, we obtain the
following approximate equation
\begin{align*}
\sum_{k=1}^m \Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)
= \sum_{k=1}^m  \Ind{v_k = \ell,y_{u_k,\ell}=-1}
\left(p_{u_k}+q_{\ell}\right)
\end{align*}

\subsection{Label propagation on $G''$}
\label{ssec:troll_lprop_weights}

Here we provide more details on the choice of weight for the edges of $G''$, as well as an
explanation on why we temporarily use symmetrized variables lying in $[-1, 1]$ (which we will denote
with primes, so that for instance $p'_u = 2p_u-1$). Since only the ratio between the negative and
positive weights matters, we fix the negative weight of the edges in $E''\setminus E'$ to be $-1$
and we denote by $\eps$ the weight of edges in $E'$.  With these notations, Label Propagation on
$G''$ seeks the harmonic minimizer of the following expression

\begin{equation*}
  \frac{1}{16} \sum_{u,v \in E} \Bigl[ \eps\left(y_{u,v} - p_u'\right)^2 + \eps\left(y_{u,v} - q_v'\right)^2 + (p'_i+q_v')^2 \Bigr]
\end{equation*}

which can be successively rewritten as
